---
title: "Intro Stat in R"
author: "No Co R Users Group"
date: "June 18, 2018"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!-- You will not be able to see this text. (i.e., this is an HTML comment) for R code use # -->

<!-- The following is just simple CSS formatting for two column tables -->
<style>
  .col2 {
    columns: 2 200px;         /* number of columns and width in pixels*/
    -webkit-columns: 2 200px; /* chrome, safari */
    -moz-columns: 2 200px;    /* firefox */
  }
  .col3 {
    columns: 3 100px;
    -webkit-columns: 3 100px;
    -moz-columns: 3 100px;
  }
</style>


## Overview
In this lab we will learn how to retrieve basic summary statistics and statistical inference from R.

## About the data
We will be using data from the "first day survey" that is administered to Stat 113 (i.e., Intro Stat) students on the first of class. (For privacy reasons, I am using data from several semesters ago...and not telling you which semester it is from!) It is available to download at: [http://myslu.stlawu.edu/~iramler/stat201/datasets/stat113.csv](http://myslu.stlawu.edu/~iramler/stat201/datasets/stat113.csv)
 

<div class="col2">
- `Year`: Class Year (of Student)
- `Gender`: Binary Gender
- `Hgt`: Height in inches
- `Wgt`: Weight in pounds
- `Smoke`: Smoking status
- `Hand`: Handedness
- `Sibs`: Number of Siblings
- `Birth`: Birth order (1 if only child)
- `MathSAT`: Math SAT score
- `VerbalSAT`: Verbal SAT score
- `GPA`: GPA (left blank for Fall semester first year students)
- `Exercise`: Hours per week
- `Greek`: Greek house status
- `Sport`: On a sports team?
- `TV`: Typical number of hours of TV (including online streaming) watched per week
- `Random`: A "randomly" chosen 4 digit number
- `Award`: Preferred Award
- `Pulse`: Pulse rate (beats per minute)
- `Pierces`: Number of piercings
- `Facebook`: (Approximate) Number of Facebook friends
- `Tattoo`: Tattoo?
- `Twitter`: Twitter user?
- `Computer`: Computer Operating System
- `SocialMedia`: Favorite social media platform
- `Class`: Section identifier

</div>


### Create a project
The first step in any new analysis that you do should be to create a project. If you haven't already done so, do it now. (Recall that you can both create a project in a new directory or by linking a project to an existing directory that already contains data, code, etc.)


### Load packages

The second step in most analyses is to load packages that you intend to use. Often, I will add extra lines to this R chunk as I build a file. Let's start by loading the readr package. (readr, part of the Tidyverse, is a useful package for loading in well-structured data. It is also what the "Import Dataset" option uses.)

```{r load-packages, message=FALSE}
library(readr)
# add more packages here as you need them
```


### Read in Data
The next step is to read in the data. (I recommend writing out the code instead of using the Import Dataset option. This helps improve reproduciblity of your code/markdown file as the Import Dataset option does not use local paths.)

```{r, eval = T, echo = T}

stat113 <- read_csv("data/stat113.csv")

```


### Understand the Data's Structure
After reading data, you often want to View() and/or glimpse() the data. (You'll likely want to make a few plots too, but that is for another day.)

```{r}
# note, don't leave a View() function active in a markdown file, 
# it won't compile correctly if you do.

dplyr::glimpse(stat113) # requires dplyr package

```

### Descriptive statistics

+ via summary()

```{r}
summary(stat113)
```


+ investigating individual variables

```{r}
mean(stat113$Hgt)
mean(stat113$GPA)
table(stat113$SocialMedia)
```


+ Correlation/Assocations


```{r}
#cor(stat113)
numericVars <- which( sapply(stat113, class) != "character")
round( cor(stat113[,numericVars], use = "p"), 2)
corrplot::corrplot(cor(stat113[,numericVars], use = "p"))

table(stat113$Sport, stat113$Award)

```


+ Any requests?


### Simple inference
**Disclaimer: I did not check the validility of the assumptions for normal-theory based inference.**

+ Construct a 95% confidence interval for the mean GPA of Stat 113 students. Also, test if the mean GPA is higher than 3.0

```{r}
t.test(x = stat113$GPA, mu = 3, alternative = "greater")
```


+ Conduct a two-sample t-test comparing the Pulse rate for athletes vs non-athletes.

```{r}
t.test(Pulse ~ Sport, data = stat113)
```


+ Is there a relationship between choice of award and tattoo status?

```{r}
AwardVsTattoo <- chisq.test(stat113$Tattoo, stat113$Award)
AwardVsTattoo
```


+ Is there a significant linear relationship between Math SAT scores and GPA?

```{r}
lmodGPA_on_MathSAT <- lm(GPA ~ MathSAT, data = stat113)
lmodGPA_on_MathSAT
summary(lmodGPA_on_MathSAT)

```



+ Any other requests?


