---
title: "R Notebooks and R Markdown"
author: "NoCoRUG"
date: "June 20, 2018"
output: html_notebook
---
## Overview

Today we'll take a quick look at using R Markdown inside of R Notebooks as a way of documenting work.

## Markdown, R Markdown, and R Notebooks

[Markdown](https://daringfireball.net/projects/markdown/) is "...a text-to-html conversion tool for web writers." [R Markdown](http://rmarkdown.rstudio.com) is a tool for using Markdown alongisde of R to combine narrative text about an analysis with the code for the analysis in order to make document the work and make it reproducable.


R Notebooks and R Markdown have several advantages:

* [Literate programming](http://literateprogramming.com/) allows you to document your work in a way that you (later) or others (now or later) can easily understand.
* Notebooks enable you to use `#commenting` to comment your code and text to narrate your analysis.
* Your end result is easily sharable.

## Let's give it a try

### About the data
We'll use some data about trees collected by my Forest Ecology students in Fall 2017.  The data are available on our [github website](https://github.com/NoCoRUG/Datasets/blob/master/trees.csv).  


### Writing code chunks

The code chunk below loads a package to deal with **https** and reads the data directly from github.

```{r}
library(RCurl) #turns on Rcurl library (which must be installed) to deal w/https
trees <-read.csv(text = getURL("https://raw.githubusercontent.com/NoCoRUG/Datasets/master/trees.csv"))

```
A cool think about R Notebooks is that you can run code and create plots right inside
of your document.

Let's add some code to take a look at our data set:
```{r}
head(trees)
```
And now let's figure out the average dbh across species:
```{r}
summary(trees$DBH)
meanDBH<-mean(trees$DBH)
meanDBH
```
We could also make a quick plot:

```{r}
library(ggplot2)
ggplot(trees, aes(x = Type, y = DBH, color = Type))+
   geom_boxplot()
```


### Formatting text
We can format text in a variety of ways using Markdown.  R Studio provides [R Markdown Cheat Sheet](https://www.rstudio.com/wp-content/uploads/2016/03/rmarkdown-cheatsheet-2.0.pdf) that can also be accessed from the R studio Help dropdown menu:
 


We can also do things like:

   * Bullet lists
    + with sublists
    
We can include **bold** items and blockquotes:

> Isn't it fun to learn about R Notebooks?

We can also print inline code such as `read.csv` if we want to. 

We can print ~subscripted~ and ^superscripted^ text or numbers. 

And much...much...more!


## Resources to learn more
The folks at RStudio produce a [R Markdown Cheat Sheet](https://www.rstudio.com/wp-content/uploads/2016/03/rmarkdown-cheatsheet-2.0.pdf)

[This article](https://support.rstudio.com/hc/en-us/articles/205368677-R-Markdown-Dynamic-Documents-for-R) from the R Studio support pages has a number of other good links.
